---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

Read the csv from the zip, and convert the date column to Date class.
```{r}
library(data.table)
#unzip('activity.zip')
data <- read.csv('activity.csv')
data$date <- as.Date(data$date)
data <- data.table(data)
```


## What is mean total number of steps taken per day?
Total number of steps per day:
```{r}
steps.per.day <- data[, .(total = sum(steps, na.rm = T)), by = date]
steps.per.day
```

Histogram of steps per day:
```{r}
library(ggplot2)
qplot(total, data = steps.per.day, geom = 'histogram', binwidth = 2000, color = I('white'))
```

Mean of total steps per day:
```{r}
mean(steps.per.day$total)
```

Median of total steps per day:
```{r}
median(steps.per.day$total)
```

## What is the average daily activity pattern?

Number of steps across average day:
```{r}
steps.per.interval <- data[, .(mean.steps = mean(steps, na.rm = T)), by = interval]
qplot(interval, mean.steps, data = steps.per.interval, geom = 'line')
```

The interval with max steps is:
```{r}
steps.per.interval[mean.steps == max(mean.steps), interval]
```

## Imputing missing values

Total number of missing values in the dataset:
```{r}
data[is.na(steps), .N]
```

Missing values will be filled with average steps for that interval:
```{r}
setkey(data, interval)
setkey(steps.per.interval, interval)
data[, steps := as.double(steps)]
data.filled <- steps.per.interval[data]
data.filled[is.na(steps), steps := mean.steps]
```

Histogram of steps per day after imputation:
```{r}
steps.per.day <- data.filled[, .(total = sum(steps, na.rm = T)), by = date]
qplot(total, data = steps.per.day, geom = 'histogram', binwidth = 2000, color = I('white'))
```

Mean of total steps per day after imputation:
```{r}
mean(steps.per.day$total)
```

Median of total steps per day after imputation:
```{r}
median(steps.per.day$total)
```

The new mean and median values are different from the values before imputation. Imputation makes the histogram more symetric.

## Are there differences in activity patterns between weekdays and weekends?

Create a new factor indicating weekdays or weekends, and show the difference in day activities:
```{r}
data[, weekday := factor(ifelse(weekdays(date) %in% c('Saturday', 'Sunday'), 'Weekend', 'Weekday'))]
steps.per.interval <- data[, .(mean.steps = mean(steps, na.rm = T)), by = .(interval, weekday)]
qplot(interval, mean.steps, data = steps.per.interval, geom = 'line') + facet_wrap(~weekday, ncol = 1)
```


